---
title: "Localization CAD_ZBP1_Rescue"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(biomaRt)
library(DESeq2)
library(cowplot)
library(tidyverse)
library(UpSetR)
library(enrichR)
library(ggpubr)
```

```{r, }
txi <- readRDS(file = "CAD_ZBP1_Rescue_txi")
t2g <- readRDS(file = "t2g")
tpms <- data.frame(txi$abundance) 
tpms <- tpms[apply(tpms, MARGIN = 1, function(x) any(x > 5)), ] 
tidy_tpms <- tpms %>% as_tibble(rownames = "ensembl_gene_id")
```

```{r, get organized for xtail}
counts <- data.frame(txi$counts) %>% 
  round() %>% 
  rownames_to_column(var="Gene") %>% 
  rowwise() %>% 
  mutate(mincount = min(c(CAD_loxP_GFP_neurite_rep1,
                          CAD_loxP_GFP_neurite_rep2,
                          CAD_loxP_GFP_neurite_rep3,
                          CAD_loxP_GFP_soma_rep1,
                          CAD_loxP_GFP_soma_rep2,
                          CAD_loxP_GFP_soma_rep3,
                          CAD_loxP_KH34_neurite_rep1,
                          CAD_loxP_KH34_neurite_rep2,
                          CAD_loxP_KH34_neurite_rep3,
                          CAD_loxP_KH34_soma_rep1,
                          CAD_loxP_KH34_soma_rep2,
                          CAD_loxP_KH34_soma_rep4,
                          CAD_loxP_ZBP1_neurite_rep1,
                          CAD_loxP_ZBP1_neurite_rep2,
                          CAD_loxP_ZBP1_neurite_rep3,
                          CAD_loxP_ZBP1_soma_rep1,
                          CAD_loxP_ZBP1_soma_rep2,
                          CAD_loxP_ZBP1_soma_rep3))) %>% 
  filter(mincount >= 10) %>% 
  dplyr::select(-mincount) %>% 
  data.frame()

rownames(counts) <- counts$Gene

##Everything below this takes quite a while.

#neurite_reads <- counts %>% select(contains("neurite")) %>% select(contains("GFP"), contains("KH34"))
#soma_reads <- counts %>% select(contains("soma")) %>% select(contains("GFP"), contains("KH34"))
#conditions <- c("GFP", "GFP", "GFP", "KH34", "KH34",  "KH34")
#xtail.results <- xtail(soma_reads, neurite_reads, conditions, bins = 10000)
#xtail.LR <- xtail.results$resultsTable %>%
#rownames_to_column(var="Gene") %>%
#rename(soma_log2FC = mRNA_log2FC, neurite_log2FC = RPF_log2FC, log2FC_LR_v1 = log2FC_TE_v1, GFP_log2LR = GFP_log2TE, KH34_log2LR = KH34_log2TE,log2FC_LR_v2 = log2FC_TE_v2, log2FC_LR_final = log2FC_TE_final, ensembl_gene_id = Gene) %>%
#left_join(select(t2g, c(ensembl_gene_id, ext_gene)), by = "ensembl_gene_id") %>%  
#rename(Gene = ext_gene) %>%
#unique()
#saveRDS(xtail.LR, file = "xtail.LR.GFPvsKH34")

```

```{r,  xtail results}
GFPvsZBP1 <-readRDS(file = "xtail.LR.ZBP1vsGFP")
GFPvsKH34 <-readRDS(file = "xtail.LR.GFPvsKH34")
ZBP1vsKH34 <-readRDS(file = "xtail.LR.ZBP1vsKH34")

```

##More LR comparisons

```{r, }

GFPvsZBP1 %>% filter(pvalue.adjust < .05) %>% nrow()

GFPvsKH34 %>% mutate(color = ifelse(pvalue.adjust < .05, "FDR < 0.05", "ns")) %>% na.omit() %>%  ggplot(aes(x = GFP_log2LR, y = KH34_log2LR, col = color, alpha = color)) + 
  geom_point() + 
  scale_color_manual(values=c("Red", "Black")) + 
  scale_alpha_manual(values=c(1, 0.1)) +
  geom_abline(intercept = 0, slope = 1) +
  scale_x_continuous(limits = c(-10, 10)) +
  scale_y_continuous(limits = c(-10, 10))

GFPvsKH34 %>% filter(pvalue.adjust < .05) %>% nrow()

ZBP1vsKH34 %>% mutate(color = ifelse(pvalue.adjust < .05, "FDR < 0.05", "ns")) %>% na.omit() %>%  ggplot(aes(x = KH34_log2LR, y = ZBP1_log2LR, col = color, alpha = color)) + 
  geom_point() + 
  scale_color_manual(values=c("Red", "Black")) + 
  scale_alpha_manual(values=c(1, 0.1)) +
  geom_abline(intercept = 0, slope = 1) +
  scale_x_continuous(limits = c(-10, 10)) +
  scale_y_continuous(limits = c(-10, 10))

ZBP1vsKH34 %>% filter(pvalue.adjust < .05) %>% nrow()

```

##Volcano Plots that I can interpret

```{r,}
GFPvsZBP1 %>% 
  mutate(color = ifelse(pvalue.adjust < .01 & ZBP1_log2LR - GFP_log2LR > 0, "FDR < 0.01", "ns")) %>%
  na.omit %>% 
  ggplot(aes(x = ZBP1_log2LR - GFP_log2LR, y = -log10(pvalue.adjust), col = color, alpha = color)) + 
  geom_point() + 
  scale_color_manual(values=c("#4DB4E3", "Black")) + 
  scale_alpha_manual(values=c(1, 0.1)) +
  scale_x_continuous(limits = c(-4, 4)) +
  ylab("-log(FDR)") +
  annotate("text", x = 2.5, y = 7.5, label = c("49 genes\nmore localized\nby ZBP1"), size = 5) +
  theme_cowplot()

GFPvsZBP1 %>% filter(pvalue.adjust<0.01) %>% mutate(ratio = ZBP1_log2LR- GFP_log2LR) %>% filter(ratio > 0) %>% nrow()
GFPvsZBP1 %>% filter(pvalue.adjust<0.01) %>% mutate(ratio = ZBP1_log2LR- GFP_log2LR) %>% filter(ratio < 0) %>% nrow()

GFPvsKH34 %>% mutate(color = ifelse(pvalue.adjust < .01 & KH34_log2LR - GFP_log2LR > 0, "FDR < 0.01", "ns")) %>% na.omit() %>% ggplot(aes(x = KH34_log2LR - GFP_log2LR, y = -log10(pvalue.adjust), col = color, alpha = color)) +
  geom_point() +
  scale_color_manual(values=c("#e37f4d", "Black")) +
  scale_alpha_manual(values=c(1, 0.1)) +
  scale_x_continuous(limits = c(-4, 4)) +
  ylab("-log(FDR)") +
  annotate("text", x = 2.5, y = 20, label = "87 genes\nmore localized\nby ΔKH34", size = 7)+
  theme_cowplot()

GFPvsKH34 %>% filter(pvalue.adjust<0.01) %>% mutate(ratio = KH34_log2LR- GFP_log2LR) %>% filter(ratio > 0) %>% nrow()
GFPvsKH34 %>% filter(pvalue.adjust<0.01) %>% mutate(ratio = KH34_log2LR- GFP_log2LR) %>% filter(ratio < 0) %>% nrow()

ZBP1vsKH34 %>% mutate(color = ifelse(pvalue.adjust < .01 & ZBP1_log2LR - KH34_log2LR > 0, "FDR<0.01", "ns")) %>% na.omit() %>% ggplot(aes(x = ZBP1_log2LR - KH34_log2LR, y = -log10(pvalue.adjust), col = color, alpha = color)) +
  geom_point() + 
  scale_color_manual(values=c("#76458E", "Black")) + 
  scale_alpha_manual(values=c(1, 0.1)) +
  scale_x_continuous(limits = c(-4, 4)) +
  ylab("-log(FDR)") +
  annotate("text", x = 2.5, y = 22, label = "60 genes\nmore localized\nby KH34", size = 5)+
  theme_cowplot()
  

ZBP1vsKH34 %>% filter(pvalue.adjust<0.01) %>% mutate(ratio = ZBP1_log2LR- KH34_log2LR) %>% filter(ratio > 0) %>% nrow()
ZBP1vsKH34 %>% filter(pvalue.adjust<0.01) %>% mutate(ratio = ZBP1_log2LR- KH34_log2LR) %>% filter(ratio < 0) %>% nrow()
```

```{r, }
mm_BAZC <- read.table("mmGenesWithACTBZipCode.txt") %>% as_tibble()

as_tibble(GFPvsZBP1) %>% 
  left_join(mm_BAZC) %>%  
  dplyr::select(ensembl_gene_id, ZBP1_log2LR, GFP_log2LR, pvalue.adjust, width, fwd,rev,ext_gene) %>%
  mutate(sig = ifelse(pvalue.adjust < 0.01, "< 0.01", "ns"),
         color = ifelse(pvalue.adjust < 0.01 & (fwd > 0 | rev > 0), "yes_sig", ifelse(pvalue.adjust < 0.01 & fwd == 0 & rev == 0, "no_sig", "ns"))) %>%
  na.omit %>% 
  ggplot(aes(x = ZBP1_log2LR - GFP_log2LR, y = -log(pvalue.adjust), col = color, alpha = color)) + 
    geom_point() + 
    scale_color_manual(values=c("Red", "Black", "Blue")) + 
    scale_alpha_manual(values=c(1, 0.1,1)) +
    scale_x_continuous(limits = c(-4, 4)) +
    ylab("-log(FDR)") +
    annotate("text", x = c(2.5,-2.5), y = c(20,20), label = c("49 genes\nmore localized\nby ZBP1", "16 genes\nmore soma\nretained"), size = 5) +
    theme_cowplot()

as_tibble(ZBP1vsKH34) %>% 
  left_join(mm_BAZC) %>%  
  dplyr::select(ensembl_gene_id, ZBP1_log2LR, KH34_log2LR, pvalue.adjust, width, fwd,rev,ext_gene) %>%
  mutate(sig = ifelse(pvalue.adjust < 0.01, "< 0.01", "ns"),
         color = ifelse(pvalue.adjust < 0.01 & (fwd > 0 | rev > 0), "yes_sig", ifelse(pvalue.adjust < 0.01 & fwd == 0 & rev == 0, "no_sig", "ns"))) %>%
  na.omit %>% 
  ggplot(aes(x = ZBP1_log2LR - KH34_log2LR, y = -log(pvalue.adjust), col = color, alpha = color)) + 
    geom_point() + 
    scale_color_manual(values=c("Red", "Black", "Blue")) + 
    scale_alpha_manual(values=c(1, 0.1,1)) +
    scale_x_continuous(limits = c(-4, 4)) +
    ylab("-log(FDR)") +
    annotate("text", x = c(2.5,-2.5), y = c(20,20), label = c("49 genes\nmore localized\nby ZBP1", "16 genes\nmore soma\nretained"), size = 5) +
    theme_cowplot()

as_tibble(GFPvsKH34) %>% 
  left_join(mm_BAZC) %>%  
  dplyr::select(ensembl_gene_id, KH34_log2LR, GFP_log2LR, pvalue.adjust, width, fwd,rev,ext_gene) %>%
  mutate(sig = ifelse(pvalue.adjust < 0.01, "< 0.01", "ns"),
         color = ifelse(pvalue.adjust < 0.01 & (fwd > 0 | rev > 0), "yes_sig", ifelse(pvalue.adjust < 0.01 & fwd == 0 & rev == 0, "no_sig", "ns"))) %>%
  na.omit %>% 
  ggplot(aes(x = KH34_log2LR - GFP_log2LR, y = -log(pvalue.adjust), col = color, alpha = color)) + 
    geom_point() + 
    scale_color_manual(values=c("Red", "Black", "Blue")) + 
    scale_alpha_manual(values=c(1, 0.1,1)) +
    scale_x_continuous(limits = c(-4, 4)) +
    ylab("-log(FDR)") +
    annotate("text", x = c(2.5,-2.5), y = c(20,20), label = c("49 genes\nmore localized\nby ZBP1", "16 genes\nmore soma\nretained"), size = 5) +
    theme_cowplot()


```

### Things that lie on the diagonal line are similarly localized by Full length ZBP1 and the ZBP1 lacking the KH34. These are not of particular interest however may be localized by other ZBP1 RBDs. 
### Above the diagonal are genes more localized specifically by the KH34 domain than any other ZBP1 RBD
### Below the diagonal are genes more soma restricted by the H34 domain than any other ZBP1

```{r, }
alldat <- ZBP1vsKH34 %>% as_tibble() %>% mutate(try = ZBP1_log2LR - KH34_log2LR) %>% dplyr::select(ensembl_gene_id, Gene, try, log2FC_LR_final,pvalue_final)
```

```{r, }
library(ggrepel)

alldat <- GFPvsZBP1 %>% as_tibble() %>% dplyr::select(ensembl_gene_id, Gene, log2FC_LR_final,pvalue.adjust) %>% dplyr::rename(ZG_LR = log2FC_LR_final, ZG_pval = pvalue.adjust) %>% mutate(ZG_LR = -ZG_LR) %>% full_join(ZBP1vsKH34) %>% dplyr::select(ensembl_gene_id, Gene, ZG_LR, ZG_pval, log2FC_LR_final, pvalue.adjust) %>% dplyr::rename(ZK_LR = log2FC_LR_final, ZK_pval = pvalue.adjust) %>% mutate(ZK_LR = -ZK_LR) %>% full_join(., GFPvsKH34) %>% dplyr::select(ensembl_gene_id, Gene, ZG_LR, ZG_pval, ZK_LR, ZK_pval, log2FC_LR_final, pvalue.adjust) %>% dplyr::rename(KG_LR = log2FC_LR_final, KG_pval = pvalue.adjust) 

z <- alldat %>% left_join(mm_BAZC) %>%
    mutate(`ACTB ZC` = ifelse(ZK_pval < 0.01 & ZG_pval < 0.01 & (fwd > 0 | rev > 0), "yes", 
                              ifelse(ZK_pval < 0.01 & ZG_pval < 0.01 & fwd == 0 & rev == 0, "no", "ns"))) %>%
    na.omit() %>%
    unique() 
   
z %>%  ggplot(aes(x = ZG_LR, y = ZK_LR)) + 
    geom_point(aes(color = `ACTB ZC`, alpha = `ACTB ZC`)) +
    theme_cowplot() + 
    geom_smooth(aes(x = ZG_LR, y = ZK_LR), method = lm, se = FALSE, inherit.aes = FALSE, col = "black") +
    stat_cor(method = "spearman", label.y = 1) + 
    geom_hline(yintercept = 0, linetype = "dashed") +
    geom_vline(xintercept = 0, linetype = "dashed") +
    scale_alpha_manual(values = c(0.25,0.01,1)) + 
    scale_color_manual(values = c("Black","Black","Red")) + 
    labs(x = "LR ZBP1 - LR GFP", y = "LR ZBP1 - LR KH34") + 
    geom_text_repel(aes(label = Gene, color = `ACTB ZC`), data = filter(z, Gene %in% c("Ddr2", "Iqgap1","Kctd10","Nin")), )


# mm_BAZC %>% filter(ensembl_gene_id %in% shared_genes$ensembl_gene_id) %>% unique() %>% View()

```

```{r, }
FLlocgenes01 <- GFPvsZBP1 %>% filter(pvalue.adjust<0.01) %>% mutate(ratio = ZBP1_log2LR- GFP_log2LR) %>% filter(ratio > 0) %>% dplyr::select(ensembl_gene_id)
trunclocgenes01 <- GFPvsKH34 %>% filter(pvalue.adjust<0.01) %>% mutate(ratio = KH34_log2LR- GFP_log2LR) %>% filter(ratio > 0) %>% dplyr::select(ensembl_gene_id)
FLvstrunclocgenes01 <- ZBP1vsKH34 %>% filter(pvalue.adjust<0.01) %>% mutate(ratio = ZBP1_log2LR- KH34_log2LR) %>% filter(ratio > 0) %>% dplyr::select(ensembl_gene_id)

FLlocgenes05 <- GFPvsZBP1 %>% filter(pvalue.adjust<0.05) %>% mutate(ratio = ZBP1_log2LR- GFP_log2LR) %>% filter(ratio > 0) %>% dplyr::select(ensembl_gene_id)
trunclocgenes05 <- GFPvsKH34 %>% filter(pvalue.adjust<0.05) %>% mutate(ratio = KH34_log2LR- GFP_log2LR) %>% filter(ratio > 0) %>% dplyr::select(ensembl_gene_id)
FLvstrunclocgenes05 <- ZBP1vsKH34 %>% filter(pvalue.adjust<0.05) %>% mutate(ratio = ZBP1_log2LR- KH34_log2LR) %>% filter(ratio > 0) %>% dplyr::select(ensembl_gene_id)

##.05/.01 = pval
## 7/3 genes are localized when Any ZBP1 construct is overexpressed
##23/25 genes are localized in ZBP over GFP and ZBP over KH34
```

## Are the same genes localized by ZBP1 and the KH34 domain?

```{r, what are the chances}
#neurite localization agreement between FLvsGFP and FLvsKH34
sum(unlist(FLlocgenes01) %in% unlist(FLvstrunclocgenes01))
phyper(25,49,nrow(GFPvsZBP1)-49,60, lower.tail = FALSE)

#soma localization agreement between FLvsGFP and FLvsKH34
s_ZG_01 <- GFPvsZBP1 %>% filter(pvalue.adjust<0.01) %>% mutate(ratio = ZBP1_log2LR- GFP_log2LR) %>% filter(ratio < 0) %>% dplyr::select(ensembl_gene_id)
s_ZG_05 <- GFPvsZBP1 %>% filter(pvalue.adjust<0.05) %>% mutate(ratio = ZBP1_log2LR- GFP_log2LR) %>% filter(ratio < 0) %>% dplyr::select(ensembl_gene_id)
s_ZK_01 <- ZBP1vsKH34 %>% filter(pvalue.adjust<0.01) %>% mutate(ratio = ZBP1_log2LR- KH34_log2LR) %>% filter(ratio < 0) %>% dplyr::select(ensembl_gene_id)
s_ZK_05 <- ZBP1vsKH34 %>% filter(pvalue.adjust<0.05) %>% mutate(ratio = ZBP1_log2LR- KH34_log2LR) %>% filter(ratio < 0) %>% dplyr::select(ensembl_gene_id)
s_KG_01 <- GFPvsKH34 %>% filter(pvalue.adjust<0.01) %>% mutate(ratio = KH34_log2LR- GFP_log2LR) %>% filter(ratio < 0) %>% dplyr::select(ensembl_gene_id)
s_KG_05 <- GFPvsKH34 %>% filter(pvalue.adjust<0.05) %>% mutate(ratio = KH34_log2LR- GFP_log2LR) %>% filter(ratio < 0) %>% dplyr::select(ensembl_gene_id)

#soma localization agreement between FLvsGFP and FLvsKH34
sum(unlist(s_ZG_01) %in% unlist(s_ZK_01))
phyper(10,16,nrow(GFPvsZBP1)-16,47, lower.tail = FALSE)

#Genes to make oligos out of:
OligoGeneList <- s_ZG_01 %>% rbind(., s_ZK_01) %>% rbind(., FLlocgenes01) %>% rbind(FLvstrunclocgenes01) %>% unlist() %>% unique()
#or should we remove genes localized by the truncation?
OligoGeneList_trunc <- s_ZG_01 %>% rbind(., s_ZK_01) %>% rbind(., FLlocgenes01) %>% rbind(FLvstrunclocgenes01) %>% anti_join(., trunclocgenes05) %>% anti_join(., s_KG_05) %>% unlist() %>% unique()

upset_dat_01 <- list(neurite_FLvG = pull(FLlocgenes01, ensembl_gene_id), neurite_FLvT = pull(FLvstrunclocgenes01, ensembl_gene_id), soma_FLvG = pull(s_ZG_01, ensembl_gene_id), soma_FLvT = pull(s_ZK_01, ensembl_gene_id))
upset_dat_05 <- list(neurite_FLvG = pull(FLlocgenes05, ensembl_gene_id), neurite_FLvT = pull(FLvstrunclocgenes05, ensembl_gene_id), soma_FLvG = pull(s_ZG_05, ensembl_gene_id), soma_FLvT = pull(s_ZK_05, ensembl_gene_id))

upset(fromList(upset_dat_01), order.by = "freq", empty.intersections = "on")

upset(fromList(upset_dat_05), order.by = "freq", empty.intersections = "on")

```

```{r, }
library(eulerr)
fit <- function(l1,l2,l3, n1, n2, n3) {
    ab = sum(l1 %in% l2)
  ac = sum(l1 %in% l3)
  bc = sum(l2 %in% l3)
  abc = length(Reduce(intersect, list(l1, l2, l3)))
  a = length(l1)-abc-ab-ac
  b = length(l2)-abc-ab-bc
  c = length(l3)-abc-bc-ac
  f <- euler(c("A"=a, "B"=b, "C"=c, "A&B"=ab, "A&C"=ac, "B&C"=bc, "A&B&C"=abc))
  f$labels <- c(n1, n2, n3)
  return(f)
}

#GFP = "#4DB481", KH34 = "#76458E", ZBP1 = "#4DB4E3"

p_lab <- paste0("p = ", signif(phyper(25,49,nrow(GFPvsZBP1)-49,60, lower.tail = FALSE),3))
v <- fit(pull(FLlocgenes01, ensembl_gene_id), pull(FLvstrunclocgenes01, ensembl_gene_id), "", "ZBP1\nlocalized", "KH34\nlocalized", "")
plot(v, quantities = TRUE, labels = v$labels, main = paste0("Significantly Localized Genes\n", p_lab), fill = c("#4DB4E3","#76458E"))

p_lab <- paste0("p = ", signif(phyper(10,16,nrow(GFPvsZBP1)-16,47, lower.tail = FALSE),3))
v <- fit(pull(s_ZG_01, ensembl_gene_id), pull(s_ZK_01, ensembl_gene_id), "", "ZBP1 soma\nretained", "KH34 soma\nretained", "")
plot(v, quantities = TRUE, labels = v$labels, main = paste0("Significantly Soma Retained Genes\n",p_lab),  fill = c("#4DB4E3","#76458E"))

p_lab <- paste0("p = ", signif(phyper(6,49,nrow(GFPvsZBP1)-49,87, lower.tail = FALSE),3))
v <- fit(pull(FLlocgenes01, ensembl_gene_id), pull(trunclocgenes01, ensembl_gene_id), "", "Full Length ZBP1\nlocalized", "Truncated ZBP1\nlocalized", "")
plot(v, quantities = TRUE, labels = v$labels, main = paste0("Significantly Localized Genes\n Genes\n",p_lab),  fill = c("#4DB4E3","#e37f4d"))

```

## Are localization targets enriched for CLIP peaks?

```{r, }
h2m <- readRDS("/Users/raegoering/Documents/TaliaferroLab/ABHaLoSeq/human2mousegeneid.txt") %>% as_tibble()

# https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4839292/
# gene-level analysis of log2 fold-enrichment scores for  eCLIP samples relative to paired size-matched input, calculated by genic region. - indicates regions not quantified as they fell below minimum read depth cutoffs (see Methods).

ipsc <- read.csv("NIHMS771565-supplement-3.csv",header = TRUE, skip = 3) %>% as_tibble()
ipsc <- ipsc %>% dplyr::select(Gene.Symbol,ENSEMBL.ID, IMP1.Rep1.CDS, IMP1.Rep1.3..UTR, IMP1.Rep2.CDS, IMP1.Rep2.3..UTR, IgG.CDS,IgG.3..UTR) %>% rowwise() %>%  mutate(ENSEMBL.ID = str_split(ENSEMBL.ID, pattern = "\\.")[[1]][1]) %>% dplyr::rename("Gene" = 1, "ensembl_gene_id" = 2, "R1_CDS" = 3, "R1_UTR" = 4, "R2_CDS" = 5, "R2_UTR" = 6,"IgG_CDS" = 7, "IgG_UTR" = 8) %>% left_join(h2m)

# 674 utr bound, 121 cds bound, only 26 bind both
ipsc <- ipsc %>% filter(IgG_UTR < 1.5, IgG_CDS <1.5) %>% mutate(UTR = ifelse(R1_UTR > 2.5 & R2_UTR > 2.5, "yes", "no"), CDS = ifelse(R1_CDS > 2.5 & R2_CDS > 2.5, "yes", "no"))
ipsc_either <- ipsc %>% filter(CDS == "yes" | UTR == "yes") %>% pull(mouse_gene)


# CLIPdb http://111.198.139.65/RBP.html
# human IGF2BP1, eclip only
postar <- read.csv("POSTAR3_CLIPdb_module_RBP_binding_sites_sub_info.csv",header = TRUE) %>% as_tibble() %>% filter(ProtocolPAR.CLIP.PARalyzerPAR.CLIP.Piranha_0.01eCLIP == "eCLIP")
postar <- postar %>% dplyr::select(1,2,3,6) %>% dplyr::rename("Gene" = 1, "ensembl_gene_id" = 2, genetype = 3, binding = 4)

#95 bound genes
postargenes <- postar %>% filter(binding > 150) %>% left_join(h2m) %>% pull(mouse_gene)

```

```{r, }
p <- GFPvsZBP1 %>% as_tibble() %>% mutate(ZBP1loc = ifelse(log2FC_LR_final < 0 & pvalue.adjust < 0.01, "yes", "no"), postar = ifelse(ensembl_gene_id %in% postargenes, "yes", "no"), ipsc = ifelse(ensembl_gene_id %in% ipsc_either, "yes","no")) %>% group_by(ZBP1loc, ipsc) %>% summarise(hits = n()) %>% spread(ipsc,hits)
p_lab <- paste0("p = ", p[2:3] %>% as.matrix() %>% fisher.test() %>% .$p.value %>% round(.,digits = 3))
p  %>% ggplot(aes(x = ZBP1loc, y = yes/no, fill = ZBP1loc)) + geom_bar(stat = "identity") + labs(x = "", y = "ZBP1 CLIP target fraction", title = "iPSC CLIP targets on ZBP1 localized genes") + theme_cowplot() + scale_x_discrete(limits = c("no", "yes"), labels = c("Localization\nUnchanged", "Localization\nIncreased")) + guides(fill = "none") + geom_text(aes(x = "no", y = 0.1, label = p_lab), hjust = -1.5) + scale_fill_manual(values = c("#eef2f5", "#4DB4E3"))


p <- GFPvsZBP1 %>% as_tibble() %>% mutate(ZBP1loc = ifelse(log2FC_LR_final < 0 & pvalue.adjust < 0.01, "yes", "no"), postar = ifelse(ensembl_gene_id %in% postargenes, "yes", "no"), ipsc = ifelse(ensembl_gene_id %in% ipsc_either, "yes","no")) %>% group_by(ZBP1loc, postar) %>% summarise(hits = n()) %>% spread(postar,hits)
p_lab <- paste0("p = ", p[2:3] %>% as.matrix() %>% fisher.test() %>% .$p.value %>% round(.,digits = 3))
p  %>% ggplot(aes(x = ZBP1loc, y = yes/no, fill = ZBP1loc)) + geom_bar(stat = "identity") + labs(x = "", y = "ZBP1 CLIP target fraction", title = "ENCODE CLIP targets on ZBP1 localized genes") + theme_cowplot() + scale_x_discrete(limits = c("no", "yes"), labels = c("Localization\nUnchanged", "Localization\nIncreased")) + guides(fill = "none") + geom_text(aes(x = "no", y = 0.075, label = p_lab), hjust = -1.5) + scale_fill_manual(values = c("#eef2f5", "#4DB4E3"))

p <- ZBP1vsKH34 %>% as_tibble() %>% mutate(ZBP1loc = ifelse(log2FC_LR_final < 0 & pvalue.adjust < 0.01, "yes", "no"), postar = ifelse(ensembl_gene_id %in% postargenes, "yes", "no"), ipsc = ifelse(ensembl_gene_id %in% ipsc_either, "yes","no")) %>% group_by(ZBP1loc, ipsc) %>% summarise(hits = n()) %>% spread(ipsc,hits)
p_lab <- paste0("p = ", p[2:3] %>% as.matrix() %>% fisher.test() %>% .$p.value %>% round(.,digits = 3))
p  %>% ggplot(aes(x = ZBP1loc, y = yes/no, fill = ZBP1loc)) + geom_bar(stat = "identity") + labs(x = "", y = "ZBP1 CLIP target fraction", title = "iPSC CLIP targets on KH34 localized genes") + theme_cowplot() + scale_x_discrete(limits = c("no", "yes"), labels = c("Localization\nUnchanged", "Localization\nIncreased")) + guides(fill = "none") + geom_text(aes(x = "no", y = 0.165, label = p_lab), hjust = -1.5) + scale_fill_manual(values = c("#eef2f5", "#76458E"))


p <- ZBP1vsKH34 %>% as_tibble() %>% mutate(ZBP1loc = ifelse(log2FC_LR_final < 0 & pvalue.adjust < 0.01, "yes", "no"), postar = ifelse(ensembl_gene_id %in% postargenes, "yes", "no"), ipsc = ifelse(ensembl_gene_id %in% ipsc_either, "yes","no")) %>% group_by(ZBP1loc, postar) %>% summarise(hits = n()) %>% spread(postar,hits)
p_lab <- paste0("p = ", p[2:3] %>% as.matrix() %>% fisher.test() %>% .$p.value %>% round(.,digits = 3))
p  %>% ggplot(aes(x = ZBP1loc, y = yes/no, fill = ZBP1loc)) + geom_bar(stat = "identity") + labs(x = "", y = "ZBP1 CLIP target fraction", title = "ENCODE CLIP targets on KH34 localized genes") + theme_cowplot() + scale_x_discrete(limits = c("no", "yes"), labels = c("Localization\nUnchanged", "Localization\nIncreased")) + guides(fill = "none") + geom_text(aes(x = "no", y = 0.075, label = p_lab), hjust = -1.5) + scale_fill_manual(values = c("#eef2f5", "#76458E"))

KH34loc <- FLlocgenes01$ensembl_gene_id[FLlocgenes01$ensembl_gene_id %in% FLvstrunclocgenes01$ensembl_gene_id]

p <- ZBP1vsKH34 %>% as_tibble() %>% mutate(ZBP1loc = ifelse(ensembl_gene_id %in% KH34loc, "yes", "no"), postar = ifelse(ensembl_gene_id %in% postargenes, "yes", "no"), ipsc = ifelse(ensembl_gene_id %in% ipsc_either, "yes","no")) %>% group_by(ZBP1loc, ipsc) %>% summarise(hits = n()) %>% spread(ipsc,hits)
p_lab <- paste0("p = ", p[2:3] %>% as.matrix() %>% fisher.test() %>% .$p.value %>% round(.,digits = 3))
p  %>% ggplot(aes(x = ZBP1loc, y = yes/no, fill = ZBP1loc)) + geom_bar(stat = "identity") + labs(x = "", y = "ZBP1 CLIP target fraction", title = "iPSC CLIP targets on shared loc hits") + theme_cowplot() + scale_x_discrete(limits = c("no", "yes"), labels = c("Localization\nUnchanged", "Localization\nIncreased")) + guides(fill = "none") + geom_text(aes(x = "no", y = 0.2, label = p_lab), hjust = -1.5) + scale_fill_manual(values = c("#eef2f5", "#5d50a2"))


p <- ZBP1vsKH34 %>% as_tibble() %>% mutate(ZBP1loc = ifelse(ensembl_gene_id %in% KH34loc, "yes", "no"), postar = ifelse(ensembl_gene_id %in% postargenes, "yes", "no"), ipsc = ifelse(ensembl_gene_id %in% ipsc_either, "yes","no")) %>% group_by(ZBP1loc, postar) %>% summarise(hits = n()) %>% spread(postar,hits)

# p[2:3] %>% as.matrix()
p_lab <- paste0("p = ns")
p  %>% ggplot(aes(x = ZBP1loc, y = yes/no, fill = ZBP1loc)) + geom_bar(stat = "identity") + labs(x = "", y = "ZBP1 CLIP target fraction", title = "ENCODE CLIP targets on shared loc hits") + theme_cowplot() + scale_x_discrete(limits = c("no", "yes"), labels = c("Localization\nUnchanged", "Localization\nIncreased")) + guides(fill = "none") + geom_text(aes(x = "no", y = 0.1, label = p_lab), hjust = -1.5) + scale_fill_manual(values = c("#eef2f5", "#5d50a2"))

##KH34trunc

p <- GFPvsKH34 %>% as_tibble() %>% mutate(KH34loc = ifelse(log2FC_LR_final < 0 & pvalue.adjust < 0.01, "yes", "no"), postar = ifelse(ensembl_gene_id %in% postargenes, "yes", "no"), ipsc = ifelse(ensembl_gene_id %in% ipsc_either, "yes","no")) %>% group_by(KH34loc, ipsc) %>% summarise(hits = n()) %>% spread(ipsc,hits)
p_lab <- paste0("p = ", p[2:3] %>% as.matrix() %>% fisher.test() %>% .$p.value %>% round(.,digits = 3))
p  %>% ggplot(aes(x = KH34loc, y = yes/no, fill = KH34loc)) + geom_bar(stat = "identity") + labs(x = "", y = "ZBP1 CLIP target fraction", title = "iPSC CLIP targets on ZBP1 localized genes") + theme_cowplot() + scale_x_discrete(limits = c("no", "yes"), labels = c("Localization\nUnchanged", "Localization\nIncreased")) + guides(fill = "none") + geom_text(aes(x = "no", y = 0.1, label = p_lab), hjust = -1.5) + scale_fill_manual(values = c("#eef2f5", "#e37f4d"))
```

```{r, }
`%notin%` <- Negate(`%in%`)

FLlocgenes01 %>% mutate(loc = "FL")
FLvstrunclocgenes01 %>% mutate(loc = "KH34")
shared_genes <- FLlocgenes01 %>% filter(ensembl_gene_id %in% FLvstrunclocgenes01$ensembl_gene_id) %>% mutate(loc = "shared")
locunchanged <- rbind(GFPvsZBP1[1:2], ZBP1vsKH34[1:2]) %>% as_tibble() %>% filter(ensembl_gene_id %notin% c(FLlocgenes01$ensembl_gene_id, FLvstrunclocgenes01$ensembl_gene_id)) %>% mutate(loc = "unchanged") %>% dplyr::select(-soma_log2FC) %>% unique()

q <- bind_rows(FLlocgenes01 %>% mutate(loc = "FL"), 
               FLvstrunclocgenes01 %>% mutate(loc = "KH34"),
               shared_genes,
               locunchanged) %>%
  mutate(postar = ifelse(ensembl_gene_id %in% postargenes, "yes", "no"),
         ipsc = ifelse(ensembl_gene_id %in% ipsc_either, "yes","no")) %>% 
  group_by(loc, ipsc) %>%
  summarise(hits = n()) %>% 
  spread(ipsc,hits)

q_lab <- q %>% mutate(matrix = list(c(yes, 595, no, 11987)),
                      fisher = fisher.test(matrix(unlist(matrix),nrow =2))$p.value,
                      label = paste0("p = ", fisher %>% round(.,digits = 3)),
                      label2 = ifelse(fisher < 0.001, "***", ifelse(fisher < 0.01, "**", ifelse(fisher < 0.05, "*", ""))))
                      
q  %>% ggplot(aes(x = loc, y = yes/no, fill = loc)) + geom_bar(stat = "identity") + labs(x = "", y = "ZBP1 CLIP target fraction", title = "iPSC CLIP targets on ZBP1 localized genes") + theme_cowplot() + scale_x_discrete(labels = c("ZBP1\nLocalized","KH34\nLocalized","Shared\nLocalization","Localization\nUnchanged")) + guides(fill = "none") + geom_text(aes(x = loc, y = yes/no + 0.01,label = label2), data = q_lab, size = 7) + scale_fill_manual(values = c("#4DB4E3","#76458E","#5d50a2", "#eef2f5"))

```

```{r, GO, fig.width = 25}
ensembl <- useMart("ensembl", dataset="mmusculus_gene_ensembl")

shared_genes_table <- getBM(attributes=c('ensembl_gene_id',
                   'external_gene_name', 'external_gene_source', 'description'),
      filters = "ensembl_gene_id",
      values = pull(shared_genes, ensembl_gene_id),
      mart = ensembl) 

```

```{r, }

dbs <- listEnrichrDbs()
dbs <- c("GO_Molecular_Function_2018", "GO_Cellular_Component_2018", "GO_Biological_Process_2018" , "ChEA_2016" ,"KEGG_2019_Mouse")

soma_gene_list <- s_ZG_05 %>% rbind(., s_ZK_05) %>% unique()
soma_gene_name <- getBM(attributes=c('ensembl_gene_id',
                   'external_gene_name'),
      filters = "ensembl_gene_id",
      values = pull(soma_gene_list, ensembl_gene_id),
      mart = ensembl) %>% pull(., external_gene_name)

neurite_gene_list <- FLlocgenes05 %>% rbind(., FLvstrunclocgenes05) %>% unique()
neurite_gene_name <- getBM(attributes = c("ensembl_gene_id", "external_gene_name"), filters = "ensembl_gene_id", values = pull(neurite_gene_list, ensembl_gene_id), mart = ensembl) %>% pull(., external_gene_name)

soma_genes <- enrichr(soma_gene_name, dbs)
soma_genes[["GO_Biological_Process_2018"]] %>% as_tibble() %>% arrange(Adjusted.P.value, P.value) %>% head(20) %>% dplyr::select(Term, Overlap, P.value, Adjusted.P.value)
soma_genes[["GO_Molecular_Function_2018"]] %>% as_tibble() %>% arrange(Adjusted.P.value, P.value) %>% head(20) %>% dplyr::select(Term, Overlap, P.value, Adjusted.P.value)
soma_genes[["GO_Cellular_Component_2018"]] %>% as_tibble() %>% arrange(Adjusted.P.value, P.value) %>% head(20) %>% dplyr::select(Term, Overlap, P.value, Adjusted.P.value)

neurite_genes <- enrichr(neurite_gene_name, dbs)
neurite_genes[["GO_Biological_Process_2018"]] %>% as_tibble() %>% arrange(Adjusted.P.value, P.value) %>% head(20) %>% dplyr::select(Term, Overlap, P.value, Adjusted.P.value)
neurite_genes[["GO_Molecular_Function_2018"]] %>% as_tibble() %>% arrange(Adjusted.P.value, P.value) %>% head(20) %>% dplyr::select(Term, Overlap, P.value, Adjusted.P.value)
neurite_genes[["GO_Cellular_Component_2018"]] %>% as_tibble() %>% arrange(Adjusted.P.value, P.value) %>% head(20) %>% dplyr::select(Term, Overlap, P.value, Adjusted.P.value)

```

## Now with TOPGO

```{r, }


GEA <- function(interesting_genes, ontology, genelist = FALSE){
 
   topDiffGenes <- function(allScore) {
   return(allScore < 0.01)
  }
  
  gene_universe_FDR <- alldat %>%  pull(ZK_pval)
  names(gene_universe_FDR) <- alldat %>% pull(ensembl_gene_id) 
  
  gene_universe_int <- factor(as.integer(names(gene_universe_FDR) %in% interesting_genes))
  names(gene_universe_int) <- names(gene_universe_FDR)
  
  go_data <- new("topGOdata",
                 ontology = ontology,
                 allGenes = gene_universe_int,
                 nodeSize = 5,
                 annotationFun = annFUN.org,
                 mapping = "org.Mm.eg",
                 ID = "ensembl")
  
  test.stat <- new("classicCount", testStatistic = GOFisherTest, name = "Fisher test")
  resultFisher <- getSigGroups(go_data, test.stat)
  allGO <- usedGO(go_data)
  if(genelist == FALSE) {
   table <- GenTable(go_data, pval = resultFisher, orderBy = "weight", ranksOf = "classic", topNodes = length(allGO)) %>% as_tibble()
    return(table)
  } 
  if(genelist == TRUE){
    return(sigGenes(go_data))
  }
  #print(knitr::kable(GenTable(go_data, pval = resultFisher, orderBy = "weight", ranksOf = "classic", topNodes = 20), caption = title))
  #showSigOfNodes(go_data, score(resultFisher), firstSigNodes = 5, useInfo = 'all')
  #return(table)
  
}
```

```{r, echo = FALSE, eval = FALSE}
`%notin%` <- Negate(`%in%`)


a <- GEA(shared_genes$ensembl_gene_id, "BP")
b <- GEA(shared_genes$ensembl_gene_id, "MF")
c <- GEA(shared_genes$ensembl_gene_id, "CC")
sh_GO <- bind_rows(list("BP" = a, "MF" = b, "CC" = c), .id = "ontology")
write.table(sh_GO, "KH34loc_geneGO.txt")


```

I have plotted the top 20 significant gene ontologies. Color indicates significance, while size of dot indicates the proportion of enriched genes within the GO.

```{r, fig.width = 10, fig.height=10, echo = FALSE}
sh_GO <- as_tibble(read.table("KH34loc_geneGO.txt", header = TRUE))

filter(sh_GO, pval < 0.05) %>% 
    arrange(pval) %>% 
    head(20) %>% 
    ggplot(aes(x = -log10(as.numeric(pval)), 
               y = factor(Term, levels = rev(unique(Term))),
               col = as.numeric(pval), 
               size = Significant/Annotated)) +
    geom_point() + 
    theme_minimal_grid() +
    labs(y = "", title = "Shared localized Gene Significant GO terms") + viridis::scale_color_viridis(begin = 0.35, direction = -1)

filter(sh_GO, pval < 0.05) %>% 
  arrange(pval) %>% 
  head(20) %>% ggplot(aes(x = -log10(pval), y = factor(Term, levels = rev(unique(Term)))), fill = -log10(pval)) + geom_bar(orientation = "y", stat = "identity", aes(fill = -log10(pval))) + theme_cowplot() + viridis::scale_fill_viridis(begin = 0.35) + labs(y = "", main = "Enriched Gene Ontologies\nZBP1 shared localized genes")
```

## SmiFISH candidates

```{r, smFISH candidate selection}

##Candidates should be genes localized by KH34 domain but not the truncated ZBP1
##Also need high enough expression (TPM > 20)


TPM_ave <- tpms %>% as_tibble(rownames = "GeneID") %>% dplyr::select(GeneID, everything()) %>% gather(-GeneID, key = sample, value = TPM) %>% separate(sample, into = c("cell", "lox","rescue","compartment","rep")) %>% group_by(GeneID, compartment) %>% summarise(ave = mean(TPM)) %>% spread(value = ave, key = compartment) %>% S4Vectors::rename("neurite" = "Ave_n_TPM", "soma" = "Ave_s_TPM")

high_enough <- TPM_ave %>% filter(Ave_n_TPM > 20 && Ave_s_TPM > 20)
high_enough_soma <- TPM_ave %>% filter(Ave_s_TPM > 20)

soma_Z_K_shared_genes <- inner_join(s_ZG_05, s_ZK_05) %>% anti_join(., s_KG_05) %>% filter(ensembl_gene_id %in% pull(high_enough_soma, GeneID))
neurite_Z_K_shared_genes <- inner_join(FLlocgenes01, FLvstrunclocgenes01) %>% anti_join(., trunclocgenes05) %>% filter(ensembl_gene_id %in% pull(high_enough, GeneID))
```

```{r,  nice plots}
##Plot where all candidates are on LR map

GFPvsZBP1 %>% mutate(color = ifelse(ensembl_gene_id %in% pull(soma_Z_K_shared_genes, ensembl_gene_id), "soma", ifelse(ensembl_gene_id %in% pull(neurite_Z_K_shared_genes, ensembl_gene_id), "neurite", "other"))) %>% na.omit %>%  ggplot(aes(x = ZBP1_log2LR - GFP_log2LR, y = -log(pvalue.adjust), col = color, alpha = color)) + 
    geom_point() + 
    scale_color_manual(values=c("Blue", "Black", "Red")) + 
    scale_alpha_manual(values=c(1, 0.1, 1)) +
    scale_x_continuous(limits = c(-4, 4)) +
    ylab("-log(FDR)") +
    theme_cowplot()

ZBP1vsKH34 %>% mutate(color = ifelse(ensembl_gene_id %in% pull(soma_Z_K_shared_genes, ensembl_gene_id), "soma", ifelse(ensembl_gene_id %in% pull(neurite_Z_K_shared_genes, ensembl_gene_id), "neurite", "other"))) %>% na.omit %>%  ggplot(aes(x = ZBP1_log2LR - KH34_log2LR, y = -log(pvalue.adjust), col = color, alpha = color)) + 
    geom_point() + 
    scale_color_manual(values=c("Blue", "Black", "Red")) + 
    scale_alpha_manual(values=c(1, 0.1, 1)) +
    scale_x_continuous(limits = c(-4, 4)) +
    ylab("-log(FDR)") +
    theme_cowplot() 

```


```{r,  picking smiFISH candidates}
##Lists of these genes
GFPvsZBP1 %>% mutate(color = ifelse(ensembl_gene_id %in% pull(soma_Z_K_shared_genes, ensembl_gene_id), "soma", ifelse(ensembl_gene_id %in% pull(neurite_Z_K_shared_genes, ensembl_gene_id), "neurite", "other"))) %>% na.omit %>% filter(color == "neurite") %>% dplyr::select(ensembl_gene_id, Gene, log2FC_LR_final, pvalue.adjust) %>% arrange(pvalue.adjust, log2FC_LR_final) %>% head(10)

ZBP1vsKH34 %>% mutate(color = ifelse(ensembl_gene_id %in% pull(soma_Z_K_shared_genes, ensembl_gene_id), "soma", ifelse(ensembl_gene_id %in% pull(neurite_Z_K_shared_genes, ensembl_gene_id), "neurite", "other"))) %>% na.omit %>% filter(color == "neurite") %>% dplyr::select(ensembl_gene_id, Gene, log2FC_LR_final, pvalue.adjust) %>% arrange(pvalue.adjust, log2FC_LR_final) %>% head(10)

sm_neurite_cand <- c("ENSMUSG00000026674", "ENSMUSG00000030536") # ,"ENSMUSG00000032826")

GFPvsZBP1 %>% mutate(color = ifelse(ensembl_gene_id %in% pull(soma_Z_K_shared_genes, ensembl_gene_id), "soma", ifelse(ensembl_gene_id %in% pull(neurite_Z_K_shared_genes, ensembl_gene_id), "neurite", "other"))) %>% na.omit %>% filter(color == "soma") %>% dplyr::select(ensembl_gene_id, Gene, log2FC_LR_final, pvalue.adjust) %>% arrange(desc(log2FC_LR_final), pvalue.adjust) %>% head(10)

ZBP1vsKH34 %>% mutate(color = ifelse(ensembl_gene_id %in% pull(soma_Z_K_shared_genes, ensembl_gene_id), "soma", ifelse(ensembl_gene_id %in% pull(neurite_Z_K_shared_genes, ensembl_gene_id), "neurite", "other"))) %>%  na.omit() %>% filter(color == "soma") %>% dplyr::select(ensembl_gene_id, Gene, log2FC_LR_final, pvalue.adjust) %>% arrange(desc(log2FC_LR_final), pvalue.adjust) %>% head(10)

#Soma candidates are lowly expressed and low LR changes
tpms %>% as_tibble() %>% mutate(GeneID = rownames(tpms)) %>% filter(GeneID == "ENSMUSG00000041794" | GeneID == "ENSMUSG00000035569" | GeneID == "ENSMUSG00000029416")

sm_soma_cand <- c("ENSMUSG00000035569", "ENSMUSG00000029416", "ENSMUSG00000041794")

controls <- c("ENSMUSG00000005732", "ENSMUSG00000026812", "ENSMUSG00000025630")

##plot the selected candidates
GZ_plot <- GFPvsZBP1 %>% mutate(color = ifelse(ensembl_gene_id %in% sm_soma_cand, "soma", ifelse(ensembl_gene_id %in% sm_neurite_cand, "neurite", ifelse(ensembl_gene_id %in% controls, "control", "other")))) %>% na.omit 
GZ_plot %>% ggplot(aes(x = ZBP1_log2LR - GFP_log2LR, y = -log(pvalue.adjust), col = color, alpha = color)) + 
    geom_point() + 
    scale_color_manual(values=c("Green", "Blue", "Black", "Red")) + 
    scale_alpha_manual(values=c(1, 1, 0.1, 1)) +
    scale_x_continuous(limits = c(-4, 4)) +
    ylab("-log(FDR)") +
    geom_text(data = subset(GZ_plot, color != "other"), aes(label = Gene), position = position_jitter(width = .5, height = 1.5)) +
    theme_cowplot()

ZK_plot <- ZBP1vsKH34 %>% mutate(color = ifelse(ensembl_gene_id %in% sm_soma_cand, "soma", ifelse(ensembl_gene_id %in% sm_neurite_cand, "neurite", ifelse(ensembl_gene_id %in% controls, "control", "other")))) %>% na.omit 
ZK_plot %>% ggplot(aes(x = ZBP1_log2LR - KH34_log2LR, y = -log(pvalue.adjust), col = color, alpha = color)) + 
    geom_point() + 
    scale_color_manual(values=c("Green", "Blue", "Black", "Red")) + 
    scale_alpha_manual(values=c(1, 1, 0.1, 1)) +
    scale_x_continuous(limits = c(-4, 4)) +
    ylab("-log(FDR)") +
    geom_text(data = subset(ZK_plot, color != "other"), aes(label = Gene), position = position_jitter(width = .5, height = 1.5)) +
    theme_cowplot()



TPM_ave %>% filter(GeneID %in% sm_soma_cand | GeneID %in% sm_neurite_cand) %>%  mutate(sm_candidate = ifelse(GeneID %in% sm_soma_cand, "soma", "neurite")) %>% left_join(ZBP1vsKH34, by = c("GeneID" = "ensembl_gene_id")) %>% dplyr::select(GeneID, Gene, sm_candidate, Ave_n_TPM, Ave_s_TPM, log2FC_LR_final, pvalue.adjust) %>% S4Vectors::rename("log2FC_LR_final" = "ZK_FC", "pvalue.adjust" = "ZK_p") %>% left_join(GFPvsZBP1, by = c("GeneID" = "ensembl_gene_id")) %>% dplyr::select(GeneID, Gene.x, sm_candidate, Ave_n_TPM, Ave_s_TPM, ZK_FC, ZK_p, log2FC_LR_final, pvalue.adjust) %>% S4Vectors::rename("Gene.x" = "Gene", "log2FC_LR_final" = "ZG_FC", "pvalue.adjust" = "ZG_p")

```

```{r, }
norm_counts <- read.table("ZBP1fractionationNormCounts.txt")

norm_counts_tidy <- norm_counts %>% as_tibble(rownames = "ensembl_gene_id") %>% gather(-ensembl_gene_id, key = sample, value = norm_counts) %>% mutate(sample = ifelse(sample == "CAD_loxP_KH34_soma_rep4", "CAD_loxP_KH34_soma_rep3", sample)) %>% separate(sample, into = c("CAD", "loxP", "construct", "fraction", "rep")) %>% dplyr::select(-CAD, -loxP) %>% left_join(., t2g[2:3])  %>% unique()  %>% spread(fraction,norm_counts)

norm_counts_tidy %>% filter(ext_gene %in% c("Myrip", "Ankrd11", "Iqgap1", "Ddr2")) %>% ggplot(aes(x = construct, y = log2(neurite/soma))) + geom_point() + facet_grid(.~factor(ext_gene,levels = c("Iqgap1", "Ddr2","Ankrd11","Myrip"))) + theme_cowplot() + scale_color_manual(values = c("#4DB481", "#76458E", "#4DB4E3")) + scale_x_discrete(limits = c("ZBP1", "KH34", "GFP")) + labs(x = "", y = "Neuronal LR, log2")

norm_counts_tidy %>% filter(ext_gene %in% c("Myrip", "Ankrd11", "Iqgap1", "Ddr2")) %>% ggplot(aes(x = construct, y = log2(neurite/soma), color = construct)) + geom_point(size = 4) + facet_grid(.~factor(ext_gene,levels = c("Iqgap1", "Ddr2","Ankrd11","Myrip"))) + theme_cowplot() + scale_color_manual(values = c("#4DB481", "#76458E", "#4DB4E3")) + scale_x_discrete(limits = c("ZBP1", "KH34", "GFP")) + labs(x = "", y = "Neuronal LR, log2") + scale_color_manual(values = c("#4DB481", "#76458E", "#4DB4E3")) + theme(strip.background = element_rect(color = "white", fill = "white")) + theme(text = element_text(size=20), axis.text = element_text(size = 12.5)) + guides(color = "none")

norm_counts_tidy %>% filter(ext_gene %in% c( "Iqgap1", "Ddr2")) %>% ggplot(aes(x = construct, y = log2(neurite/soma), color = construct)) + geom_point(size = 4) + facet_grid(.~ext_gene) + theme_cowplot() + scale_color_manual(values = c("#4DB481", "#76458E", "#4DB4E3")) + scale_x_discrete(limits = c("ZBP1", "KH34", "GFP")) + labs(x = "", y = "Neuronal LR, log2") + theme(strip.background = element_rect(color = "white", fill = "white")) + theme(text = element_text(size=20), axis.text = element_text(size = 12.5)) + guides(color = "none")

```

```{r, making something helpful I hope}
#rbind(FL_loc_seq, trunc_loc_seq) %>% rbind(., FL_ctrl_seq) %>% rbind(., trunc_ctrl_seq) %>% rbind(., FLvT_loc_seq) %>% rbind(FLvT_ctrl_seq) %>% filter(ensembl_gene_id %in% OligoGeneList_trunc) %>% unique() -> oligo_gene_seq

#regex <- oligo_gene_seq %>% rowwise() %>%  mutate(sub = list(allsubstr(seq, nchar("ACACCCACACCC"))), min_BAZC = length(agrep("ACACCCACACCC", unlist(sub), max.distance = 2)),  BPZC = str_count(string = seq, pattern = BAZC_BP)) %>% select(ensembl_gene_id, min_BAZC, BPZC)

# GFPvsZBP1 %>% filter(ensembl_gene_id %in% OligoGeneList_trunc) %>% as_tibble() %>% dplyr::select(ensembl_gene_id, Gene, log2FC_LR_final,pvalue_final) %>% dplyr::rename(ZG_LR = log2FC_LR_final, ZG_pval = pvalue_final) %>% left_join(ZBP1vsKH34) %>% dplyr::select(ensembl_gene_id, Gene, ZG_LR, ZG_pval, log2FC_LR_final, pvalue_final) %>% dplyr::rename(ZK_LR = log2FC_LR_final, ZK_pval = pvalue_final) %>% left_join(., GFPvsKH34) %>% dplyr::select(ensembl_gene_id, Gene, ZG_LR, ZG_pval, ZK_LR, ZK_pval, log2FC_LR_final, pvalue_final) %>% dplyr::rename(KG_LR = log2FC_LR_final, KG_pval = pvalue_final) %>% left_join(TPM_ave, by = c("ensembl_gene_id" = "GeneID")) %>% left_join(., regex) -> savethis


#saveRDS(savethis, file = "ZBP1_Oligo_gene_list_with_info_no_controls.txt")
```

## 3'UTR selection

```{r, UTRclone candidates}
#ddr2 iqgap1, myo6, bicd2
ZCpos <- c("ENSMUSG00000026674", "ENSMUSG00000030536", "ENSMUSG00000033577", "ENSMUSG00000037933")
#ank2, kctd10, nin, ppp1r9a
ZCneg <- c("ENSMUSG00000032826", "ENSMUSG00000001098", "ENSMUSG00000021068", "ENSMUSG00000032827 ")

GZ_plot <- GFPvsZBP1 %>% mutate(actbZC = ifelse(ensembl_gene_id %in% ZCpos, "pos", ifelse(ensembl_gene_id %in% ZCneg, "neg", "other"))) %>% na.omit 
GZ_plot %>% ggplot(aes(x = ZBP1_log2LR - GFP_log2LR, y = -log(pvalue.adjust), col = actbZC, alpha = actbZC)) + 
    geom_point() + 
    scale_color_manual(values=c("Blue", "Black", "Red")) + 
    scale_alpha_manual(values=c(1, 0.1, 1)) +
    scale_x_continuous(limits = c(-4, 4)) +
    ylab("-log(FDR)") +
    geom_text(data = subset(GZ_plot, actbZC != "other"), aes(label = Gene), position = position_jitter(width = .5, height = 1.5)) +
    theme_cowplot()

ZK_plot <- ZBP1vsKH34 %>% mutate(actbZC = ifelse(ensembl_gene_id %in% ZCpos, "pos", ifelse(ensembl_gene_id %in% ZCneg, "neg", "other"))) %>% na.omit 
ZK_plot %>% ggplot(aes(x = ZBP1_log2LR - KH34_log2LR, y = -log(pvalue.adjust), col = actbZC, alpha = actbZC)) + 
    geom_point() + 
    scale_color_manual(values=c("Blue", "Black", "Red")) + 
    scale_alpha_manual(values=c(1, 0.1, 1)) +
    scale_x_continuous(limits = c(-4, 4)) +
    ylab("-log(FDR)") +
    geom_text(data = subset(ZK_plot, actbZC != "other"), aes(label = Gene), position = position_jitter(width = .5, height = 1.5)) +
    theme_cowplot()

```

