---
title: "CAD_ZBP1_Rescue"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tximport)
library(biomaRt)
library(stringr)
library(ggfortify)
library(DESeq2)
library(pheatmap)
library(ggpubr)
library(tidyverse)
library(cowplot)
```

```{r, tximport stuff}
mart <- useMart("ENSEMBL_MART_ENSEMBL",
                dataset = "mmusculus_gene_ensembl",
                host='www.ensembl.org')

t2g <- getBM(attributes = c('ensembl_transcript_id',
                            'ensembl_gene_id',
                            'external_gene_name',
                            'refseq_mrna'), mart = mart)

t2g <- rename(t2g, target_id = ensembl_transcript_id, 
              ext_gene = external_gene_name)

ens2gene <- t2g[,c(2,3)]
colnames(ens2gene)[2] <- 'Gene'
ens2gene <- unique(ens2gene)

base_dir <- "CAD_ZBP1_Rescue_quants"

files <- list.files(base_dir)
files <- files[c(1:9,11:12,14:20)]
samples <- str_sub(files, 1, str_length(files)-9)

salm_dirs <- sapply(samples, function(id) file.path(paste(base_dir,"/",id,".quant.sf",sep = "")))
tx2gene <- t2g[,c(1,2)]
colnames(tx2gene) <- c('TXNAME', 'GENEID')

txi <- tximport(salm_dirs, 
                type = 'salmon', 
                tx2gene = tx2gene,
                dropInfReps = TRUE, 
                countsFromAbundance = 'lengthScaledTPM')

tpms <- data.frame(txi$abundance) 
tpms <- tpms[apply(tpms, MARGIN = 1, function(x) any(x > 5)), ] 

```

###A quick look at TPMs

```{r,  looknsee}
dat <- t(tpms) %>% 
  as_tibble() %>% 
  mutate(sample = colnames(tpms))

dat_all <- dat %>%
  separate(sample, into = c("cell", "loxp", "treatment", "fraction", "rep"), sep = "_") %>% 
  dplyr::select(treatment, fraction, rep, everything(), -cell, -loxp)

autoplot(prcomp(log(dat_all[4:ncol(dat_all)]+1)), 
         data = dat_all,
         colour = "treatment", 
         shape = "fraction",
         size = 5) +
  ggtitle("PCA of log(TPMs)") +
  theme_cowplot() +
  scale_color_manual(values = c("#4DB481", "#76458E", "#4DB4E3"))

```

###Now for normalized counts

```{r,  DEseq stuff}
conditions <- c("GFP_neurite", "GFP_neurite", "GFP_neurite", "GFP_soma", "GFP_soma", "GFP_soma", "KH34_neurite", "KH34_neurite", "KH34_neurite", "KH34_soma", "KH34_soma", "KH34_soma", "ZBP1_neurite", "ZBP1_neurite", "ZBP1_neurite", "ZBP1_soma", "ZBP1_soma", "ZBP1_soma")
seq_run <- c("1", "1", "1", "1", "1", "1", "1", "1", "1", "1", "1", "2", "1", "1", "1", "1", "1", "1")
type <- c("paired-end", "paired-end", "paired-end", "paired-end", "paired-end", "paired-end", "paired-end", "paired-end", "paired-end", "paired-end","paired-end", "paired-end", "paired-end", "paired-end", "paired-end", "paired-end", "paired-end", "paired-end")
colData <- data.frame("sample" = samples, "Conditions" = conditions, "seq_run" = seq_run, "Type" = type)

dds <- DESeqDataSetFromTximport(txi, colData = colData, design = ~seq_run + Conditions)
dds <- DESeq(dds)

norm_counts <- counts(dds,normalized = TRUE)
norm_counts <- norm_counts[apply(norm_counts, MARGIN = 1, function(x) any(x > 100)), ]

norm_counts_PCAdat <- t(norm_counts) %>%
  as_tibble() %>%
  mutate(sample = colnames(norm_counts)) %>%
  separate(sample, into = c("cell", "loxp", "treatment", "fraction", "rep"), sep = "_") %>% 
  dplyr::select(treatment, fraction, rep, everything(), -cell, -loxp)

autoplot(prcomp(log(norm_counts_PCAdat[4:ncol(norm_counts_PCAdat)]+1)),
         data = norm_counts_PCAdat,
         colour = "treatment",
         size = 5, 
         shape = "fraction") +
  ggtitle("PCA of log(normalized counts)") +
  theme_cowplot() +
  scale_color_manual(values = c("#4DB481", "#76458E", "#4DB4E3"))

```


###More normalized counts

```{r,  more DEseq stuff}

temp <- norm_counts
colnames(temp) <- substr(colnames(norm_counts),10, 28)
annoDF <- as.data.frame(colnames(temp)) %>% as_tibble() %>% separate("colnames(temp)",into = c("Construct", "Fraction", "rep"), sep = "_" ) %>% dplyr::select(Construct, Fraction) %>% as.data.frame()

ann_colors = list(
    Fraction = c(soma = "white", neurite = "black"),
    Construct = c(GFP = "#4DB481", KH34 = "#76458E", ZBP1 = "#4DB4E3"))

rownames(annoDF) <- colnames(temp)
cor_mat <- temp %>% cor(method = "spearman")

pheatmap(cor_mat, annotation_col = annoDF, annotation_colors = ann_colors)

dds2 <- DESeq(dds,test="LRT",reduced = ~1)

AllChanges <- results(dds2)
rlog <- rlog(dds2)
rlogMatrix <- assay(rlog)

sigChanges <- rownames(AllChanges)[AllChanges$padj < 0.001 & !is.na(AllChanges$padj)]
sigMat <- rlogMatrix[rownames(rlogMatrix) %in% sigChanges,]

annoDF <- as.data.frame(colData(rlog)[,1,drop=FALSE]) %>%
  as_tibble(rownames = "rowname") %>%  
  mutate(sample = str_sub(sample, 10, str_length(sample)-5)) %>%
  as.data.frame()
rownames(annoDF) <- annoDF$rowname
annoDF <- annoDF[2]

pheatmap(sigMat,
         show_rownames = FALSE,
         annotation_col = annoDF,
         main = "heatmap of log(norm counts)") 

```

##What about PCA separation on localization ratios?

```{r, LR from norm counts}
LR_tidy <- norm_counts %>%
  as.data.frame() %>%  
  as_tibble(rownames = "Gene") %>%
  mutate(LR_GFP_rep1 = CAD_loxP_GFP_neurite_rep1/CAD_loxP_GFP_soma_rep1, LR_GFP_rep2 = CAD_loxP_GFP_neurite_rep2/CAD_loxP_GFP_soma_rep2, LR_GFP_rep3 = CAD_loxP_GFP_neurite_rep3/CAD_loxP_GFP_soma_rep3, LR_KH34_rep1 = CAD_loxP_KH34_neurite_rep1/CAD_loxP_KH34_soma_rep1, LR_KH34_rep2 = CAD_loxP_KH34_neurite_rep2/CAD_loxP_KH34_soma_rep2, LR_KH34_rep3 = CAD_loxP_KH34_neurite_rep3/CAD_loxP_KH34_soma_rep4, LR_ZBP1_rep1 = CAD_loxP_ZBP1_neurite_rep1/CAD_loxP_ZBP1_soma_rep1, LR_ZBP1_rep2 = CAD_loxP_ZBP1_neurite_rep2/CAD_loxP_ZBP1_soma_rep2, LR_ZBP1_rep3 = CAD_loxP_ZBP1_neurite_rep3/CAD_loxP_ZBP1_soma_rep3) %>% 
  dplyr::select(Gene, contains("LR"))

is.na(LR_tidy)<-sapply(LR_tidy, is.infinite)
LR_tidy[is.na(LR_tidy)] <- 0

LRPCAdat <- LR_tidy %>% 
  dplyr::select(-Gene) %>% 
  as.data.frame()

rownames(LRPCAdat) <- LR_tidy$Gene
LRPCAdat <- t(LRPCAdat) %>% 
  as_tibble() %>%
  mutate(sample = colnames(LRPCAdat)) %>%
  separate(sample, into = c("LR", "treatment", "rep"), sep = "_") %>%
  dplyr::select(treatment, rep, everything(), -LR)

autoplot(prcomp(log2(LRPCAdat[3:ncol(LRPCAdat)]+1)),
         data = LRPCAdat, 
         colour = "treatment",
         size = 5) +
  ggtitle("PCA of log2(Localization Ratio)") +
  theme_cowplot() +
  guides(size = FALSE)


```

```{r, LR from norm counts no outlier}

cor_mat <- log2(LR_tidy[2:10]) %>% cor(method = "spearman")
pheatmap(cor_mat)

sigLR <- LR_tidy[apply(LR_tidy[, -1], MARGIN = 1, function(x) any(abs(x) > 2)), ] %>% as.data.frame()
rownames(sigLR) <- sigLR$Gene

annoDF <- c(LRPCAdat$treatment) %>% as.data.frame()
rownames(annoDF) <- colnames(sigLR)[2:10]
colnames(annoDF) <- "Sample"

pheatmap(sigLR[2:10],
         show_rownames = FALSE,
         annotation_col = annoDF,
         scale = "row",
         main = "heatmap of log2(LR)") 

```


##Now I'll stop trying to interpret PCA plots and try to determine the quality of the fractionation 

###How do the fractions differ in TPM of different genes

```{r, }
tidy_dat <- norm_counts %>%
  as_tibble() %>% 
  mutate(ensembl_gene_id = rownames(norm_counts)) %>%
  left_join(t2g[,2:3]) %>%
  rename("gene" = ext_gene) %>% 
  unique() %>% 
  gather(key = sample, value = norm_counts, -gene, - ensembl_gene_id) %>%
  separate(sample, into = c("cell", "loxp", "treatment", "fraction", "rep"), sep = "_") %>% 
  mutate(norm_counts = as.numeric(norm_counts)) %>% 
  dplyr::select(gene, treatment, fraction, rep, norm_counts)

tidy_dat %>% filter(gene == "Igf2bp1") %>% ggplot(aes(x = treatment, y = as.numeric(norm_counts), fill = fraction)) + geom_boxplot() + geom_point(position = position_jitterdodge()) + labs(x = "", y = "Normalized Counts", title = "ZBP1") +
  theme_cowplot()
tidy_dat %>% filter(gene == "Actb") %>% ggplot(aes(x = treatment, y = as.numeric(norm_counts), fill = fraction)) + geom_boxplot() + geom_point(position = position_jitterdodge()) + labs(x = "", y = "Normalized Counts", title = "Beta Actin") + theme_cowplot()

tidy_dat %>% filter(gene == "Malat1") %>% ggplot(aes(x = treatment, y = as.numeric(norm_counts), fill = fraction)) + geom_boxplot() + geom_point(position = position_jitterdodge()) + labs(x = "", y = "Normalized Counts", title = "Malat1") + theme_cowplot()

tidy_dat %>% filter(gene == "Gapdh") %>% ggplot(aes(x = treatment, y = as.numeric(norm_counts), fill = fraction)) + geom_boxplot() + geom_point(position = position_jitterdodge()) + labs(x = "", y = "Normalized Counts", title = "Gapdh") + theme_cowplot()

tidy_dat %>% filter(gene == "Tsc1") %>% ggplot(aes(x = treatment, y = as.numeric(norm_counts), fill = fraction)) + geom_boxplot() + geom_point(position = position_jitterdodge()) + labs(x = "", y = "Normalized Counts", title = "Tsc1") + theme_cowplot()

tidy_dat %>% filter(gene == "Ranbp1") %>% ggplot(aes(x = treatment, y = as.numeric(norm_counts), fill = fraction)) + geom_boxplot() + geom_point(position = position_jitterdodge()) + labs(x = "", y = "Normalized Counts", title = "Ranbp1") + theme_cowplot()

saveRDS(tidy_dat, file = "CAD_ZBP1_Rescue_tpms")
saveRDS(txi, file = "CAD_ZBP1_Rescue_txi")
saveRDS(t2g, file = "t2g")
```

###How do the fractions differ in LR of different genes

```{r, }
LR_tidier <- LR_tidy %>%
  left_join(t2g[,2:3], by = c("Gene" = "ensembl_gene_id")) %>%
  rename("gene" = ext_gene) %>% 
  unique() %>% 
  gather(key = sample, value = LR, -gene, -Gene) %>%
  separate(sample, into = c("LRlabel", "treatment", "rep"), sep = "_") %>% 
  mutate(LR = as.numeric(LR)) %>% dplyr::select(gene, treatment, rep, LR)

LR_tidier %>% filter(gene == "Igf2bp1") %>% ggplot(aes(x = treatment, y = as.numeric(log2(LR)), col = treatment, size = 5)) + geom_point() + labs(x = "", y = "Log2(Localization Ratio)", title = "ZBP1") + guides(col = FALSE, size = FALSE) + theme_cowplot()

LR_tidier %>% filter(gene == "Actb") %>% ggplot(aes(x = treatment, y = as.numeric(log2(LR)), col = treatment, size = 5)) + geom_point() + labs(x = "", y = "Log2(Localization Ratio)", title = "Beta Actin") + guides(col = FALSE, size = FALSE) + theme_cowplot()
LR_tidier %>% filter(gene == "Malat1") %>% ggplot(aes(x = treatment, y = as.numeric(log2(LR)), col = treatment, size = 5))+ geom_point() + labs(x = "", y = "Log2(Localization Ratio)", title = "Malat1") + guides(col = FALSE, size = FALSE) + theme_cowplot()
LR_tidier %>% filter(gene == "Gapdh") %>% ggplot(aes(x = treatment, y = as.numeric(log2(LR)), col = treatment, size = 5)) + geom_point() + labs(x = "", y = "Log2(Localization Ratio)", title = "Gapdh") + guides(col = FALSE, size = FALSE) + theme_cowplot()

saveRDS(LR_tidier, file = "CAD_ZBP1_Rescue_LR")
```

###comparing the LRs

```{r, }
LR_tidier %>% group_by(gene, treatment) %>% summarise(mean = mean(LR)) %>% spread(key = treatment, value = mean) %>% ggplot(aes(x = log2(GFP), y = log2(ZBP1))) + geom_point() + geom_abline(intercept = 0, slope = 1) + theme_cowplot()
LR_tidier %>% group_by(gene, treatment) %>% summarise(mean = mean(LR)) %>% spread(key = treatment, value = mean) %>% ggplot(aes(x = log2(GFP), y = log2(KH34))) + geom_point() + geom_abline(intercept = 0, slope = 1) + theme_cowplot()
LR_tidier %>% group_by(gene, treatment) %>% summarise(mean = mean(LR)) %>% spread(key = treatment, value = mean) %>% ggplot(aes(x = log2(ZBP1), y = log2(KH34))) + geom_point() + geom_abline(intercept = 0, slope = 1) + theme_cowplot()

```

##Now a DEseq object comparing all neurite to all soma

```{r, }
conditions <- c("neurite", "neurite", "neurite", "soma", "soma", "soma", "neurite", "neurite", "neurite", "soma", "soma", "soma", "neurite", "neurite", "neurite", "soma", "soma", "soma")
seq_run <- c("1", "1", "1", "1", "1", "1", "1", "1", "1", "1", "1", "2", "1", "1", "1", "1", "1", "1")
type <- c("paired-end", "paired-end", "paired-end", "paired-end", "paired-end", "paired-end", "paired-end", "paired-end", "paired-end", "paired-end", "paired-end","paired-end", "paired-end", "paired-end", "paired-end", "paired-end", "paired-end", "paired-end")
colData <- data.frame("sample" = samples, "Conditions" = conditions, "seq_run" = seq_run, "Type" = type)

dds_qc <- DESeqDataSetFromTximport(txi, colData = colData, design = ~Conditions)
dds_qc <- DESeq(dds_qc)
res <- results(dds_qc, contrast = c("Conditions", "neurite", "soma"))

```

###some GO subsetting

```{r, }
simpleres <- res %>%
  as_tibble(rownames = "ensembl_gene_id") %>%
  dplyr::select(ensembl_gene_id, log2FoldChange) %>% 
  mutate(GO = "All Genes") %>%
  left_join(t2g[2:3]) %>% 
  unique()

#Ribosomegenes <- getBM(attributes = c('ensembl_gene_id'), filters = c('go_parent_term'), values = c('GO:0005840'), mart = mart) %>% as_tibble() %>% mutate(GO = "Ribosome") %>% left_join(.,simpleres, by = "ensembl_gene_id") %>% select(ensembl_gene_id,log2FoldChange,GO.x) %>% rename("GO" = "GO.x")

#ETCgenes <- getBM(attributes = c('ensembl_gene_id'), filters = c('go_parent_term'), values = c('GO:0022900'), mart = mart) %>% as_tibble() %>% mutate(GO = "ETC") %>% left_join(.,simpleres, by = "ensembl_gene_id") %>% select(ensembl_gene_id,log2FoldChange,GO.x) %>% rename("GO" = "GO.x")

#nucleargenes <- getBM(attributes = c('ensembl_gene_id'), filters = c('go_parent_term'), values = c('GO:0005634'), mart = mart) %>% as_tibble() %>% mutate(GO = "nuclear") %>% left_join(.,simpleres, by = "ensembl_gene_id") %>% select(ensembl_gene_id,log2FoldChange,GO.x) %>% rename("GO" = "GO.x")

#synapsegenes <- getBM(attributes = c('ensembl_gene_id'), filters = c('go_parent_term'), values = c('GO:0045202'), mart = mart) %>% as_tibble() %>% mutate(GO = "synapse") %>% left_join(.,simpleres, by = "ensembl_gene_id") %>% select(ensembl_gene_id,log2FoldChange,GO.x) %>% rename("GO" = "GO.x")

#saveRDS(Ribosomegenes, file = "Ribosomal_Protein_Gene_list")
#saveRDS(ETCgenes, file = "ETC_Gene_list")
#saveRDS(nucleargenes, file = "nuclear_Gene_list")
#saveRDS(synapsegenes, file = "synapse_Gene_list")

Ribosomegenes <-readRDS(file = "Ribosomal_Protein_Gene_list") %>% left_join(t2g[2:3]) %>% unique()
ETCgenes <-readRDS(file = "ETC_Gene_list") %>% left_join(t2g[2:3]) %>% unique()
nucleargenes <-readRDS(file = "nuclear_Gene_list") %>% left_join(t2g[2:3]) %>% unique()
synapsegenes <- readRDS(file = "synapse_Gene_List") %>% left_join(t2g[2:3]) %>% unique()

ETCgenes_mm <- readRDS(file = "/Users/raegoering/Documents/TaliaferroLab/ABHaLoSeq/mm_ETCgenes.txt")
Mito_Mat_mm <- readRDS(file = "/Users/raegoering/Documents/TaliaferroLab/ABHaLoSeq/mm_mitomatrixgenes.txt")
Mito_tx_mm <- readRDS(file = "/Users/raegoering/Documents/TaliaferroLab/ABHaLoSeq/mm_mitotranscript.txt")

RPgenes <- simpleres %>% filter(startsWith(prefix = "Rpl", ext_gene) | startsWith(prefix = "Rps", ext_gene)) %>% mutate(GO = "RP")
ETCgenes2 <- simpleres %>% filter(ensembl_gene_id %in% ETCgenes_mm) %>% mutate(GO = "ETC2")
Mito_Matgenes <- simpleres %>% filter(ensembl_gene_id %in% Mito_Mat_mm) %>% mutate(GO = "Mito_Mat")
Mito_txgenes <- simpleres %>% filter(ensembl_gene_id %in% Mito_tx_mm) %>% mutate(GO = "Mito_tx")


lnc <- getBM(attributes= c('ensembl_gene_id'),filters = "biotype", value = 'lncRNA', mart = mart)
pro <- getBM(attributes= c('ensembl_gene_id'),filters = "biotype", value = 'protein_coding',mart = mart)
sno <- getBM(attributes= c('ensembl_gene_id'),filters = "biotype", value = 'snoRNA',mart = mart)
sn <- getBM(attributes= c('ensembl_gene_id'),filters = "biotype", value = 'snRNA',mart = mart)


lncgenes <- simpleres %>% filter(ensembl_gene_id %in% lnc$ensembl_gene_id) %>% mutate(GO = "lnc")
progenes <- simpleres %>% filter(ensembl_gene_id %in% pro$ensembl_gene_id) %>% mutate(GO = "pro")
snogenes <- simpleres %>% filter(ensembl_gene_id %in% sno$ensembl_gene_id) %>% mutate(GO = "sno")
sngenes <- simpleres %>% filter(ensembl_gene_id %in% sn$ensembl_gene_id) %>% mutate(GO = "sn")


```

```{r, }

my_comparisons <- list(c("All Genes", "ETC"),
                    c("All Genes", "nuclear"),
                    c("All Genes", "Ribosome"),
                    c("All Genes", "synapse"),
                    c("All Genes", "ETC2"),
                    c("All Genes", "Mito_Mat"),
                    c("All Genes", "Mito_tx"),
                    c("All Genes", "RP"),
                    c("All Genes", "pro"),
                    c("All Genes", "lnc"),
                    c("All Genes", "sno"),
                    c("All Genes", "sn"))

rbind(simpleres,Ribosomegenes,ETCgenes,nucleargenes,synapsegenes,RPgenes,Mito_Matgenes,Mito_txgenes,ETCgenes2,progenes,lncgenes,snogenes,sngenes) %>%
    as_tibble() %>% na.omit() %>% 
    ggplot(aes(x = GO, y = log2FoldChange, fill = GO)) +
    geom_boxplot(notch = TRUE, outlier.shape = NA)  +
    geom_hline(yintercept = 0, linetype = "dashed") +
    EnvStats::stat_n_text(y.pos = -1.5) +
    stat_compare_means(comparisons = my_comparisons, method = "wilcox.test",label = "p.signif", size = 7, label.y = c(1.0, 1.25, 1.5)) +
    coord_cartesian(ylim = c(-2,3)) + 
    guides(fill = FALSE) +
    ylab("log2(Localization Ratio)") +
    guides(fill = FALSE) +
    theme_cowplot()

###pretty version

my_comparisons <- list(c("All Genes", "pro"),
                       c("All Genes", "sno"),
                       c("All Genes", "ETC2"),
                       c("All Genes", "RP"))

rbind(simpleres,RPgenes,ETCgenes2,progenes,snogenes) %>%
    as_tibble() %>% na.omit() %>% 
    ggplot(aes(x = GO, y = log2FoldChange, fill = GO)) +
    geom_violin() +
    geom_boxplot(outlier.shape = NA, width = 0.3)  +
    geom_hline(yintercept = 0, linetype = "dashed") +
    EnvStats::stat_n_text(y.pos = -1.8) +
    stat_compare_means(comparisons = my_comparisons, method = "wilcox.test",label = "p.signif", size = 7, vjust = 0.5, label.y = c(0.5,0.75,1.0,1.25)) +
    coord_cartesian(ylim = c(-2,3.5)) + 
    guides(fill = FALSE) +
    ylab("log2(Localization Ratio)") +
    guides(fill = FALSE) +
    theme_cowplot() +
    scale_x_discrete(limits = c("All Genes", "pro","sno","ETC2","RP"), labels = c("All Genes", "mRNA", "snoRNA","ETC","RP mRNA")) +
    scale_fill_manual(values = c("#c90076","#ce7e00","#8fce00","#2986cc","#6a329f")) +
    labs(x = "", y = "Neuronal Localization Ratio, log2")


```

```{r, }
library(topGO)

topDiffGenes <- function(allScore) {
   return(allScore < 0.01)
}

# all genes
GEA_all <- function(ontology){
  gene_universe_FDR <- res %>% as_tibble(rownames = "ensembl_gene_id") %>% pull(padj)
  names(gene_universe_FDR) <- res %>% as_tibble(rownames = "ensembl_gene_id") %>% pull(ensembl_gene_id)
  
  go_data <- new("topGOdata",
                   ontology = ontology,
                   allGenes = gene_universe_FDR,
                   geneSel = topDiffGenes,
                   nodeSize = 5,
                   annotationFun = annFUN.org,
                   mapping = "org.Hs.eg",
                   ID = "ensembl")
    
    test.stat <- new("classicCount", testStatistic = GOFisherTest, name = "Fisher test")
    resultFisher <- getSigGroups(go_data, test.stat)
  allGO <- usedGO(go_data)
  GenTable(go_data, pval = resultFisher, orderBy = "weight", ranksOf = "classic", topNodes = length(allGO))
}

GEA <- function(interesting_genes, ontology, genelist = FALSE){
 
   topDiffGenes <- function(allScore) {
   return(allScore < 0.01)
  }
  
  gene_universe_FDR <- res %>% as_tibble(rownames = "ensembl_gene_id") %>% pull(padj)
  names(gene_universe_FDR) <- res %>% as_tibble(rownames = "ensembl_gene_id") %>% pull(ensembl_gene_id)
  

  gene_universe_int <- factor(as.integer(names(gene_universe_FDR) %in% interesting_genes))
  names(gene_universe_int) <- names(gene_universe_FDR)
  
  go_data <- new("topGOdata",
                 ontology = ontology,
                 allGenes = gene_universe_int,
                 nodeSize = 5,
                 annotationFun = annFUN.org,
                 mapping = "org.Mm.eg",
                 ID = "ensembl")
  
  test.stat <- new("classicCount", testStatistic = GOFisherTest, name = "Fisher test")
  resultFisher <- getSigGroups(go_data, test.stat)
  allGO <- usedGO(go_data)
  if(genelist == FALSE) {
   table <- GenTable(go_data, pval = resultFisher, orderBy = "weight", ranksOf = "classic", topNodes = length(allGO)) %>% as_tibble()
    return(table)
  } 
  if(genelist == TRUE){
    return(sigGenes(go_data))
  }
  #print(knitr::kable(GenTable(go_data, pval = resultFisher, orderBy = "weight", ranksOf = "classic", topNodes = 20), caption = title))
  #showSigOfNodes(go_data, score(resultFisher), firstSigNodes = 5, useInfo = 'all')
  #return(table)
  
}


```

```{r, eval = FALSE}
Ngenes <- res %>% as_tibble(rownames = "ensembl_gene_id") %>% filter(padj < 0.05, log2FoldChange > 1.5) %>% pull(ensembl_gene_id)
Sgenes <- res %>% as_tibble(rownames = "ensembl_gene_id") %>% filter(padj < 0.05, log2FoldChange < -1.5) %>% pull(ensembl_gene_id)

pos_BP <- GEA(Ngenes, "BP")
pos_MF <- GEA(Ngenes, "MF")
pos_CC <- GEA(Ngenes, "CC")
pos_GO <- bind_rows(list("BP" = pos_BP, "MF" = pos_MF, "CC" = pos_CC), .id = "ontology")
write.table(pos_GO, "ZBP1_Neurite_localized_GO.txt")

neg_BP <- GEA(Sgenes, "BP")
neg_MF <- GEA(Sgenes, "MF")
neg_CC <- GEA(Sgenes, "CC")
neg_GO <- bind_rows(list("BP" = neg_BP, "MF" = neg_MF, "CC" = neg_CC), .id = "ontology")
write.table(neg_GO, "ZBP1_Soma_localized_GO.txt")

```

```{r, echo = FALSE}
N_GO <- as_tibble(read.table("ZBP1_Neurite_localized_GO.txt", header = TRUE))
S_GO <- as_tibble(read.table("ZBP1_Soma_localized_GO.txt", header = TRUE))
```

```{r, fig.width = 10, fig.height=10, echo = FALSE}
N_GO %>% 
  mutate(pval=ifelse(grepl(pattern = "<", pval), 
                     substr(pval,3,nchar(pval)), pval), 
         pval = as.numeric(pval)) %>% 
  filter(pval < 0.00001) %>% 
  arrange(pval) %>% 
  head(20) %>% 
  ggplot(aes(x = -log(pval), 
             y = factor(Term, levels = rev(Term)),
             col = pval, 
             size = Significant/Annotated)) +
  geom_point() + 
  theme_minimal_grid() +
  labs(y = "")

S_GO %>% 
  mutate(pval=ifelse(grepl(pattern = "<", pval), 
                     substr(pval,3,nchar(pval)), pval), 
         pval = as.numeric(pval)) %>% 
  filter(pval < 0.00001) %>% 
  arrange(pval) %>% 
  head(20) %>% 
  ggplot(aes(x = -log(pval), 
             y = factor(Term, levels = rev(Term)),
             col = pval, 
             size = Significant/Annotated)) +
  geom_point() + 
  theme_minimal_grid() +
  labs(y = "")
```


```{r, }
enst_txi <- tximport(salm_dirs, 
                type = 'salmon', 
                tx2gene = tx2gene,
                dropInfReps = TRUE, 
                txOut = TRUE,
                countsFromAbundance = 'lengthScaledTPM')

enst_txi <- data.frame(enst_txi$abundance) 

```